A decomposition Approach for solving the JSP using multi-shot ASP\thanks{Supported by organization x.}

### 2022-04-12 ICLP Paper/samplepaper.tex
checked experiments section

In order to counteract limitations of window-wise successive optimization
due to "decomposition mistakes" and unnecessary idle times that do not directly
affect the makespan to some extent,
we have devised two additional techniques that can be incorporated into the
iterative scheduling process.
The first extension is time window \emph{overlapping}, where a configurable percentage
of the operations per time window can still be rescheduled when proceeding
to the next time window.
To this end, an (optimized) partial schedule is postprocessed and the configured
number of operations to overlap are picked in decreasing order of starting times.
For example, if one operation from the first time window is supposed overlap for the
decomposed schedule in Figure~\ref{fig:decomposed}, the second operation of job~\lstinline{2}
whose starting time~\lstinline{4} is latest may be chosen,
which in turn enables its processing after the second operation of job~\lstinline{3}
by rescheduling together with operations of the second time window.
The addition of overlapping operations to the encoding in Listing~\ref{prg:encoding}
requires handling them similar to operations of the time window to schedule,
i.e., enabling the choice of an execution order by the rules in lines
\ref{prg:base:order}-\ref{prg:base:order:end} rather than fixing the order
by the rule in line~\ref{prg:base:ordered}.
As the second extension, we can make use of a stratified ASP program to
postprocess an (optimized) partial schedule by inspecting operations of the
latest time window in the order of starting times whether idle slots
on their machines allow for an earlier execution.
The starting times obtained by the \emph{compression} are then taken instead
of those calculated in the actual schedule optimization, and, e.g.,
the starting time~\lstinline{12} of the third operation of job~\lstinline{1}
would be turned into~\lstinline{9} to fill the idle slot available on
machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}.

\section{Experiments}\label{sec:experiments}
% This section will show the computational experiments that we conducted to test the performance of our proposed model.
In order to evaluate our multi-shot ASP modulo DL approach to JSP solving,
we ran experiments on JSP benchmark sets by \cite{taillard1993benchmarks}
of several sizes,
considering ten instances each with
$50\times15$, $50\times20$, or $100\times20$ jobs and machines. \hl{We have selected the largest set of instances generated and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
The instances are generated such that each job includes one operation per machine, e.g.,
$15$ operations in case of the $50\times15$ instances,
where the sequence in which the operations of a job are allocated to machines varies.
In our experiments, we assess decomposition strategies presented in Section~\ref{subsec:decomposition}
with different number and size of time windows as well as the impact of
time window overlapping and compression techniques described in Section~\ref{subsec:scheduling}.
For the comparability of results between runs with a different number of time windows
and respective optimization subproblems addressed by multi-shot solving,
we divide the total runtime limit of $1000$ seconds for \clingodl (version 1.3.0) by the number
of time windows to evenly spend optimization efforts on subproblems.
Our experiments have been run on an 
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.%
\footnote{The benchmarks and our implementation are available at: \myurl}

@@

% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment. Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment. The results of this table conclude that the most effective decomposition strategies are machine-based regardless of whether the operations are ordered based on EST or MTWR; however, the dynamic assignment has no positive effect than the static. %The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
% We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible. Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap. The first column refers to the overlap percentage where $0\%$ means no overlap between TWs. Every three columns have presented the results of a particular set of instances. We applied the overlap and compression with the solutions of the number of windows, which provided a better average makespan using a machine-based EST strategy. In the non-compression part, the overlap has improved the results slightly through the different sizes where $10\%$ and $20\%$ of overlap provided the best results. However, the compression has shown a significant improvement wrt the running time and the average makespan, especially for the $100\times20$ set. The compression demonstrated effectiveness as it minimized the average makespan and decreased the running time compared to the model without compression. Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
\hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR tools. Table~{\ref{tab:Table04}} summarizes the results of our experiments. The values in the table represent the average distance to the optimal solutions for each set of instances. The first column represents the size of the instances with the decomposition parameters. The second and third columns show the average distance of the results obtained by ASP single and multi-shot, respectively. The last column shows the average distance obtained by OR tools. The results show that the single-shot ASP is far from the optima, especially with the largest dataset. In addition, it is shown that the developed model improves the results compared to the pure ASP. However, the performance of OR tools is better than ours, especially for the first two sets of instances. For $100 \times 20$, our proposed model is a bit comparable to OR tools, which means that if the instance size increases, it could be possible to obtain better results with our approach.}

@@

\begin{table}[t]
  \caption{Experiments comparing the Single, Multi-shot ASP and OR tools\label{tab:Table04}}%
  \centering
  \begin{tabular}{l l r r r r r r}

  \topline%\hline
  \multirow{3}{*}{Instances} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\

  & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot \\

  $50\times15$ & $2733.8$ & $2778.8$ & $3542.1$ & $2932.9$ & $5.0$ & $768.3$ & $159.1$ \\
  $50\times20$ & $2843.9$ & $2914.3$ & $3879.4$ & $3120.3$ & $70.4$ & $1035.5$ & $276.4$ \\
  $100\times20$ & $5365.7$ & $5559.3$ & $46786.1$ & $5637.5$ & $193.6$ & $41420.4$ & $271.8$

  \botline %\hline
  \end{tabular}
\end{table}

\section{Conclusions}\label{sec:conclusions}
%
Our work develops multi-shot ASP modulo DL methods for JSP solving,
based on problem decomposition into balanced time windows that respect the operation
precedence within jobs and successive makespan optimization while extending
partial schedules to a global solution.
We demonstrate that splitting highly complex JSP instances into
balanced portions, for which partial schedules of good quality can be reliably
found within tight runtime limits, leads to significantly better global solutions
than obtainable with single-shot ASP modulo DL optimization in similar runtime.
Related works taking advantage of single- \citep{abjoossctowa21a}
and multi-shot solving \citep{francescutto2021solving} by
\clingodl to tackle real-world scheduling problems evaluate threshold values
on DL variables and perform optimization by means of common 
\lstinline{#minimize} statements of \clingo rather than minimizing the
value of a DL variable like \lstinline{makespan} in Listing~\ref{prg:encoding}.
Applying a similar approach can be advantageous in our future work as it may
potentially improve computational efficiency as well as solution quality further
and is even a necessity to switch from the makespan as optimization objective to tardiness
w.r.t.\ deadlines, which are crucial in many application areas. \hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems. However, the proposed model must be changed to solve open-shop since the precedence constraint is not applied}
Moreover, we aim to generalize our methods further, e.g.,
by considering flexible machine allocations and varying processing times,
and scale them up to large-scale instances with an order of magnitude more
operations than in synthetic benchmarks.
%
% In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints. The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan. The proposed model has been developed to improve the performance by applying the overlap between TWs to fix the decomposition mistakes. In addition, the compression of each window is to fulfill the gaps that occurred during the optimization process. We tested our model on a set of benchmark instances and the experiments have shown that i) considering the bottleneck machines during the decomposition process is effective, ii) The overlap between TWs improves the solution quality, and iii) the compression of TWs is super practical, as it aided the proposed model to obtain better solutions in less computational time. Our work can be extended by linking the optimization model with a simulator that considers more realistic factors such as machine breakdowns, new job arrivals and order cancellations.

>>>>

In order to counteract limitations of window-wise successive optimization
due to ``decomposition mistakes'' and unnecessary idle times that do not directly
affect the makespan to some extent,
we have devised two additional techniques that can be incorporated into the
iterative scheduling process.
The first extension is time window \emph{overlapping}, where a configurable percentage
of the operations per time window can still be rescheduled when proceeding
to the next time window.
To this end, an (optimized) partial schedule is postprocessed and the configured
number of operations to overlap are picked in decreasing order of starting times.
For example, if one operation from the first time window is supposed overlap for the
decomposed schedule in Figure~\ref{fig:decomposed}, the second operation of job~\lstinline{2}
whose starting time~\lstinline{4} is latest may be chosen,
which in turn enables its processing after the second operation of job~\lstinline{3}
by rescheduling together with operations of the second time window.
The addition of overlapping operations to the encoding in Listing~\ref{prg:encoding}
requires handling them similar to operations of the time window to schedule,
i.e., enabling the choice of an execution order by the rules in lines
\ref{prg:base:order}-\ref{prg:base:order:end} rather than fixing the order
by the rule in line~\ref{prg:base:ordered}.
As the second extension, we can make use of a stratified ASP program to
postprocess an (optimized) partial schedule by inspecting operations of the
latest time window in the order of starting times whether idle slots
on their machines allow for an earlier execution.
The starting times obtained by the \emph{compression} are then taken instead
of those calculated in the actual schedule optimization, and, e.g.,
the starting time~\lstinline{12} of the third operation of job~\lstinline{1}
would be turned into~\lstinline{9} to fill the idle slot available on
machine~\lstinline{3} when compressing the decomposed schedule displayed in Figure~\ref{fig:decomposed}.

\section{Experiments}\label{sec:experiments}
% This section will show the computational experiments that we conducted to test the performance of our proposed model.
In order to evaluate our multi-shot ASP modulo DL approach to JSP solving,
we ran experiments on JSP benchmark sets by \cite{taillard1993benchmarks}
of several sizes,
considering ten instances each with
$50\times15$, $50\times20$, or $100\times20$ jobs and machines.
\hl{These are the largest instances provided by common JSP benchmark sets,
so that complete optimization methods do not terminate within tight runtime limits
and the amount of operations makes problem decomposition worthwhile.}
% and excluded the small instances because they could be solved by the existing methods and applying decomposition leads to unnecessary overheads.}
The instances are generated such that each job includes one operation per machine, e.g.,
$15$ operations in case of the $50\times15$ instances,
where the sequence in which the operations of a job are allocated to machines varies.
In our experiments, we assess decomposition strategies presented in Section~\ref{subsec:decomposition}
with different number and size of time windows as well as the impact of
time window overlapping and compression techniques described in Section~\ref{subsec:scheduling}.
For the comparability of results between runs with a different number of time windows
and respective optimization subproblems addressed by multi-shot solving,
we divide the total runtime limit of $1000$ seconds for \clingodl (version 1.3.0) by the number
of time windows to evenly spend optimization efforts on subproblems.
Our experiments have been run on an 
Intel\textsuperscript{\textregistered} Core\texttrademark{} i7-8650U CPU % \textsuperscript{\textregistered}
Dell Latitude 5590 machine under Windows 10.%
\footnote{The benchmarks and our implementation are available at: \myurl}

@@

\begin{table}[b]
  \caption{Comparison of single- and multi-shot ASP modulo DL solving approaches to OR-tools\label{tab:Table04}}%
  \centering
  \begin{tabular}{r r r r r r r r}

  \topline%\hline
%  \multirow{3}{*}{Size} & \multicolumn{4}{c}{Makespan} & \multicolumn{3}{c}{Distance measures} \\
%  & Optima & OR-tools & \multicolumn{2}{c}{ASP} & OR-tools & \multicolumn{2}{c}{ASP} \\

 % & & & Single-shot & Multi-shot &  & Single-shot & Multi-shot 
 & & \multicolumn{2}{c}{OR-tools} & \multicolumn{2}{c}{Multi-shot} & \multicolumn{2}{c}{Single-shot} \\
 Size & Optima & Makespan & Distance & Makespan & Distance & Makespan & Distance \midline

  $50\times15$ & $2773.8$ & $2778.8$ & $5.0$ & $2932.9$ & $159.1$ & $3542.1$ & $768.3$ \\
  $50\times20$ & $2843.9$ & $2914.3$ & $70.4$ & $3120.3$ & $276.4$ & $3879.4$ & $1035.5$ \\
  $100\times20$ & $5365.7$ & $5559.3$ & $193.6$ & $5637.5$ & $271.8$ & $46786.1$ & $41420.4$

  \botline %\hline
  \end{tabular}
\end{table}
% Table \ref{tab:Table02} demonstrates the results of the four static decomposition strategies and the machine-based MTWR with the dynamic assignment. Every three sub-columns have illustrated the results of a set of the benchmark instances and each row refers to a decomposition strategy with static (S) or dynamic (D) assignment. The results of this table conclude that the most effective decomposition strategies are machine-based regardless of whether the operations are ordered based on EST or MTWR; however, the dynamic assignment has no positive effect than the static. %The best average makespan of $50\times15$ and $100\times20$ are  $3060.3$, and $5948.6$.
% We applied the overlap between the windows to repair the mistakes generated by the decomposition; in addition, we compressed each optimized window to ensure that all the operations had been started as early as possible. Table \ref{tab:Table03} summarizes the results with/without the compression and the overlap. The first column refers to the overlap percentage where $0\%$ means no overlap between TWs. Every three columns have presented the results of a particular set of instances. We applied the overlap and compression with the solutions of the number of windows, which provided a better average makespan using a machine-based EST strategy. In the non-compression part, the overlap has improved the results slightly through the different sizes where $10\%$ and $20\%$ of overlap provided the best results. However, the compression has shown a significant improvement wrt the running time and the average makespan, especially for the $100\times20$ set. The compression demonstrated effectiveness as it minimized the average makespan and decreased the running time compared to the model without compression. Therefore, the overlap and compression have improved the makespan results by $4.5\%, 3.4\%$ and $6.8\%$ and the running time by $40\%$.
\hl{For putting our results, obtained with single-shot ASP modulo DL solving ($1$ time window in Table}~\ref{tab:Table01}\hl{) and the best-performing multi-shot solving approach
indicated in the lower half of Table}~\ref{tab:Table03}\hl{, in relation,
we also ran the state-of-the-art CP system OR-tools (version 9.2)} \citep{ortools}\hl{
with a timeout of $1000$ seconds,
using the JSP encoding provided by} \cite{tagesc21a}\hl{.
The respective average makespans and distances to the known optima reported by}
\cite{taillard1993benchmarks}\hl{ are shown in Table}~\ref{tab:Table04}\hl{.
Notably, the optimal schedules from the literature have been determined by greedy or
local search techniques, which can in general not compute (provably) optimal solutions,
yet the optimality of their schedules is guaranteed by theoretical lower bounds that were
not supplied as background knowledge to} \clingodl \hl{and OR-tools.
As a consequence, all runs of the complete OR-tools and} \clingodl \hl{solvers are
interrupted at the runtime limit and do thus not yield optimality proofs.
However, we observe that the schedules of OR-tools are of high quality and
come close to the optima for the instances of size $50\times15$.
The larger the instances get, the closer is our decomposition and multi-shot solving approach,
while it remains behind OR-tools.
On the one hand, this can be due to ``decomposition mistakes'' of our method,
deteriorating the quality of the resulting global solutions.
On the other hand, the CP encoding taken for OR-tools features dedicated interval variables
and global constraints, which are tailored to scheduling problems and give it an advantage
over the use of DL constraints.
Nevertheless, the improvements due to problem decomposition and multi-shot solving are apparent
relative to the single-shot solving approach, which is significantly behind for each of the three problem sizes.}
% \hl{We conducted other experiments to assess the quality of the solutions obtained by our proposed method by comparing ASP with single-shot and OR-tools (version 9.2)} \citep{ortools}\hl{. Table}~\ref{tab:Table04} \hl{summarizes the results of our experiments. The values in the table represent the average distance to the optimal solutions for each set of instances. The first column represents the size of the instances with the decomposition parameters. The second and third columns show the average distance of the results obtained by ASP single and multi-shot, respectively. The last column shows the average distance obtained by OR tools. The results show that the single-shot ASP is far from the optima, especially with the largest dataset. In addition, it is shown that the developed model improves the results compared to the pure ASP. However, the performance of OR tools is better than ours, especially for the first two sets of instances. For $100 \times 20$, our proposed model is a bit comparable to OR tools, which means that if the instance size increases, it could be possible to obtain better results with our approach.}

@@

\section{Conclusions}\label{sec:conclusions}
%
Our work develops multi-shot ASP modulo DL methods for JSP solving,
based on problem decomposition into balanced time windows that respect the operation
precedence within jobs and successive makespan optimization while extending
partial schedules to a global solution.
We demonstrate that splitting highly complex JSP instances into
balanced portions, for which partial schedules of good quality can be reliably
found within tight runtime limits, leads to significantly better global solutions
than obtainable with single-shot ASP modulo DL optimization in similar runtime.
Related works taking advantage of single- \citep{abjoossctowa21a}
and multi-shot solving \citep{francescutto2021solving} by
\clingodl to tackle real-world scheduling problems evaluate threshold values
on DL variables and perform optimization by means of common 
\lstinline{#minimize} statements of \clingo rather than minimizing the
value of a DL variable like \lstinline{makespan} in Listing~\ref{prg:encoding}.
Applying a similar approach can be advantageous in our future work as it may
potentially improve computational efficiency as well as solution quality further
and is even a necessity to switch from the makespan as optimization objective to tardiness
w.r.t.\ deadlines, which are crucial in many application areas.
While Machine-based decomposition turned out to be helpful
for finding better schedules, the strategies presented in Section~\ref{subsec:decomposition}
can be regarded as naive heuristics based on straightforward instance properties,
and incorporating machine-learning approaches \citep{elscge22a,tagesc21a}
for more adept partitioning is another promising direction for future work.
\hl{Furthermore, it would be beneficial to apply the same approach to Flow-shop and open-shop scheduling problems. However, the proposed model must be changed to solve open-shop since the precedence constraint is not applied}
Moreover, we aim to generalize our methods further, e.g.,
by considering flexible machine allocations and varying processing times,
and scale them up to large real-world scheduling problems with an order of magnitude more
operations than in synthetic benchmarks.
%
% In this work, different decomposition strategies have been proposed for solving the Job-shop Scheduling Problem (JSP) using a multi-shot Answer Set Programming with Difference Logic constraints to handle the decomposition approach and the problem constraints. The proposed model split the problem into sub-problems using different decomposition strategies and then optimized each separately while minimizing the Makespan. The proposed model has been developed to improve the performance by applying the overlap between TWs to fix the decomposition mistakes. In addition, the compression of each window is to fulfill the gaps that occurred during the optimization process. We tested our model on a set of benchmark instances and the experiments have shown that i) considering the bottleneck machines during the decomposition process is effective, ii) The overlap between TWs improves the solution quality, and iii) the compression of TWs is super practical, as it aided the proposed model to obtain better solutions in less computational time. Our work can be extended by linking the optimization model with a simulator that considers more realistic factors such as machine breakdowns, new job arrivals and order cancellations.
